---
title: "r-week-8-hw-eeb177"
author: "Matthew Muia"
date: "March 4, 2017"
output: html_document
---

## Part 1: Canids Continued
# Report on output
# 1. What taxonomic group is displayed?
# The group is Cormocyon.
# 2. How many unique fossil species are contained within it?
# There are two species: copei and haydeni.
# 3. How many unique fossil genera?
# There are 58 unique fossil genera within the file. 
# 4. What species has the longest record?
# Cormocyon copei has the longest record.


## Part 2: Data Frames
# Exploring Data Frames
# Exercise #1
```{r}
df <- data.frame(first = c('Matthew'),
                 last = c('Muia'),
                 lucky_number = c(513),
                 stringsAsFactors = FALSE)
df <- rbind(df, list('Kelsey', 'Bowers', 213) )
df <- cbind(df, coffeetime = c(TRUE,F))
df
```

# Exercise #2
```{r}
download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv", destfile = "data/gapminder-FiveYearData.csv")
gapminder <- read.csv(file = "data/gapminder-FiveYearData.csv")
source(file = "scripts/load-gapminder.R")
```

# Exercise #3
The object gapminder is a data frame with columns
- country and continent are factors.
- year is an integer vector.
- pop, lifeExp, and gdpPercap are numeric vectors.

# Subsetting Data
# Exercise #1
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
x[2:4]
x[-c(1,5)]
x[c(2,3,4)]
```

# Exercise #2
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
# a and c are correct
```

# Exercise #3
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
x_subset <- x[x<7 & x>4]
print(x_subset)
```

# Exercise #4
```{r}
m <- matrix(1:18, nrow=3, ncol=6)
print(m)
# answer D
m[2,c(4,5)]
```

# Exercise #5
```{r}
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
xlist$b[2]
```

# Exercise #6
```{r}
mod <- aov(pop ~ lifeExp, data=gapminder)
attributes(mod)
mod$df.residual
```

# Exercise #7
```{r}
#1
gapminder[gapminder$year == 1957,]

#2
gapminder[,-c(1:4)]

#3
gapminder[gapminder$lifeExp > 80,]

#4
gapminder[1, c(4, 5)]

#5
gapminder[gapminder$year == 2002 | gapminder$year == 2007,]
 gapminder[gapminder$year %in% c(2002, 2007),]

```

# Exercise #8
```{r}
#1
# gapminder is a data.frame so needs to be subsetted on two dimensions. gapminder[1:20, ] subsets the data to give the first 20 rows and all columns.

#2
gapminder_small <- gapminder[c(1:9, 19:23),]
```

# Creating Publication-Quality Graphs
# Exercise #1
```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()
```

# Exercise #2
```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp, color=continent)) + geom_point()
```

# Exercise #3
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) + geom_point() + geom_line(aes(color=continent))
```

# Exercise #4
```{r}
# part a
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point(size=3, color="orange") + scale_x_log10() + geom_smooth(method="lm", size=1.5)

# part b
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + geom_point(size=3, pch=17) + scale_x_log10() + geom_smooth(method="lm", size=1.5)
```

# Exercise #5
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, fill=continent)) + geom_density(alpha=0.6) + facet_wrap( ~ year) + scale_x_log10()
```

